\large Neural and Symbolic Optimization of Cosmological Particle-Mesh Simulation \vspace{0.3em}

### 2024-03-01 docs/papers/sto/so.tex
WIP draft: figure on density slab

\begin{figure*}
  \centering
  \includegraphics[width=.48\columnwidth]{net_f_a1_j3091776_e3000.pdf}
  \includegraphics[width=.48\columnwidth]{net_g_a1_j3091776_e3000.pdf}
  \caption{$f(k_i; \vvartheta)$ and $g(k_1, k_2, k_3; \vvartheta)$
  functions. The solid line and shaded region indicate the median and
  the [16, 84] percentiles of the 64 Sobol simulation setups. $a=1$,
  mesh shape=1}
  \label{fig:func}
\end{figure*}

For both $f$ and $g$ NNs, we adopt multilayer perceptron (MLP, i.e.
fully connected) structures to learn the nonlinear functions.
Both $f$ and $g$ networks have $XXX$ hidden layers with $XXX\times n_{\rm
input}$ nodes for each layer, where $n_{\rm input}$ is the size of the
input layer, which is [...] and [...] for $f$ and $g$ respectively.
The rectified linear unit (ReLU) activation function is used for hidden
layers.
We further apply the exponential function to the output as a regulator
for positive output.
Being embedded in the differentiable simulation library \pmwd, the NNs
can be trained using gradient-based optimization methods as in usual
deep learning tasks.
The parameters of the NNs, which enter the force evaluation in every
time integration step, can be treated just the same as other
cosmological parameters.
The objective gradients are accumulated, based on which the parameters
are then optimized.

@@

Improvements of SO on the dark matter density fields are the first
things to check.
We construct these fields by scattering the particles onto a mesh.
To compare two fields $\hat f$ and $f$, we introduce for each mode the
square root ratio of the auto power spectrum
%
\begin{equation}
T(\vk) \triangleq
\sqrt{\frac{|\hat f(\vk)|^2}{|f(\vk)|^2}},
\label{T}
\end{equation}
%
which is referred to as the transfer function (TF), and the correlation
coefficient (CC) defined with the cross power spectrum
%
\begin{equation}
r(\vk) \triangleq
\frac{\Re \bigl[ \hat f(\vk) f^*(\vk) \bigr]}
     {\sqrt{|\hat f(\vk)|^2 |f(\vk)|^2}}.
\label{r}
\end{equation}
%
The two fields are identical when both $T$ and $r$ are unity.
The PM method is accurate on scales much larger than the mesh cell size,
therefore we expect $T$ and $r$ to be very close to being unity both
before and after the SO with force sharpening.

@@

\begin{figure*}
  \centering
  \includegraphics[width=.8\columnwidth]{denslab_0_120_e300.pdf}
  \caption{Projected dark matter density fields of \GADGET, \pmwd\ without and
  with force sharpening.
  Sobol 0 snap 120}
  \label{fig:denslab}
\end{figure*}

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=.48\columnwidth]{net_f_a1_j3091776_e3000.pdf}
  \includegraphics[width=.48\columnwidth]{net_g_a1_j3091776_e3000.pdf}
  \caption{$f(k_i; \vvartheta)$ and $g(k_1, k_2, k_3; \vvartheta)$
  functions. The solid line and shaded region indicate the median and
  the [16, 84] percentiles of the 64 sampled configurations and
  cosmological parameters. $a=1$,
  mesh shape=1}
  \label{fig:func}
\end{figure*}

For both the $f$ and $g$ spatial optimization NNs (SONNs), we adopt
multilayer perceptron (MLP, i.e. fully connected) structures to learn
the nonlinear functions.
Both $f$ and $g$ networks have $XXX$ hidden layers with $XXX\times n_{\rm
input}$ nodes for each layer, where $n_{\rm input}$ is the size of the
input layer, which is [...] and [...] for $f$ and $g$ respectively.
The rectified linear unit (ReLU) activation function is used for hidden
layers.
We further apply the exponential function to the output as a regulator
for positive output.
Being embedded in the differentiable simulation library \pmwd, the NNs
can be trained using gradient-based optimization methods as in usual
deep learning tasks.
The parameters of the NNs, which enter the force evaluation in every
time integration step, can be treated just the same as other
cosmological parameters.
The objective gradients are accumulated, based on which the parameters
are then optimized.

@@

Improvements of SO on the dark matter density fields are the first
things to check.
We construct these fields by scattering the particles onto a mesh.
To compare two fields $\hat f$ and $f$, we introduce for each mode the
transfer function (TF), which is defined as the square root ratio of the
auto power spectrum:
%
\begin{equation}
T(\vk) \triangleq
\sqrt{\frac{|\hat f(\vk)|^2}{|f(\vk)|^2}},
\label{T}
\end{equation}
%
and the correlation coefficient (CC):
%
\begin{equation}
r(\vk) \triangleq
\frac{\Re \bigl[ \hat f(\vk) f^*(\vk) \bigr]}
     {\sqrt{|\hat f(\vk)|^2 |f(\vk)|^2}},
\label{r}
\end{equation}
%
where the numerator is the real part of the cross power spectrum.
The two fields are identical when both $T$ and $r$ are unity.
The PM method is accurate on scales much larger than the mesh cell size,
therefore we expect $T$ and $r$ to be very close to being unity both
before and after the SO with force sharpening.

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\columnwidth]{slab_s0_snap120_3177874_e3000.pdf}
  \caption{Projected dark matter density fields of \GADGET, original PM,
  and optimized PM-SONN.
  The examples here are shown for the 0th data point in our samples of
  configurations and cosmological parameters.
  This corresponds to an overall simulation box size of 79 Mpc$/h$.
  We show a zoom-in view with window length XXX Mpc$/h$, and a projected
  slab thickness of XXX Mpc$/h$.}
  \label{fig:denslab}
\end{figure*}
